No Replication Without Process Documentation:  An Assessment of Management

### 2019-09-19 MSE manuscript.Rmd
Edit r code snippet names and break points; Remove roles, open meetings, and adopted from the freq.data analysis; create random sample only freq table and part.plot

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html")
##### load libraries #####
library(tidyverse) # upgrade to base R
library(readxl) # read in excel files
library(shiny) # Shiny App
library(shinythemes) # web themes for Shiny
library(DT) # Data Table
library(purrr) # for loop alternative
library(ggmap) # map plotting
library(mapdata) # basemap generation
library(odbc) # database connection 
library(RMySQL) # MySQL scripting in R
library(DBI) # database interface
library(kableExtra) # extra formating of the tables

@@

# Move comment column to the end
data<-data %>%
  select(-'Comments', 'Comments')

# Filter data
data_pub<-filter(data,IncludeInPublication==TRUE)
data_CC<-filter(data,str_detect(Drivers,"Climate Change"))

@@

#####-- Data analysis --#####
# Get columns for study summary
summary.col<-c("Citation",
               "Species",
               "Location",
               "System")
# Get columns for study drivers and problem
prob.col<-c("Citation",
            "ProblemDefinition",
            "Drivers",
            "ConsequencePrediction",
            "TradeOffMethod_Exp",
            "TradeOffMethod_Sub",
            "Decision")
# Get columns for frequency analysis
freq.col<-c("ProcessExplicit",
            "ProblemDefinitionExplicit",
            "ObjectivesExplicit",
            "AlternativesExplicit",
            "TradeOffsExplicit",
            "DecisionExplicit",
            "RolesExplicit",
            "OpenMeetings",
            "ResultsAdopted")
# Get columns for participant analysis
part.col<-c("Leader",
            "Participants",
            "ObjElicitationSource_Exp",
            "ProcedureElicitation_Exp",
            "ObjElicitationSource_Sub",
            "ProcedureElicitation_Sub")
obj.col<-c("ObjType",
           "ObjCategory",
           "ObjDirection",
           "ObjScale")
alt.col<-c("ManagementType",
           "AlternativesEvaluated")
# Get columns for map
map.col<-c("Latitude",
           "Longitude",
           "Citation",
           "Drivers")

# Select data summary columns
summary.data_pub<-data_pub %>%
    select(summary.col)
summary.data_CC<-data_CC %>%
  select(summary.col)

@@

freq.data_pub<-data_pub %>%
    select(freq.col) %>%
    rename("Process"="ProcessExplicit",
           "Problem"="ProblemDefinitionExplicit",
           "Objectives"="ObjectivesExplicit",
           "Alternatives"="AlternativesExplicit",
           "Tradeoffs"="TradeOffsExplicit",
           "Decision"="DecisionExplicit",
           "Roles"="RolesExplicit",
           "Open Meetings"="OpenMeetings",
           "Adopted"="ResultsAdopted") %>%
    summarise_all(funs(sum))%>%
    gather(Explicit) %>%
    mutate(Percent=value/n_mse*100) %>%
    mutate(Explicit=factor(Explicit,levels=
                  c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision",
                               "Roles","Open Meetings","Adopted"))) %>%
    rename("Number"="value")

freq.data_CC<-data_CC %>%
  select(freq.col) %>%
  rename("Process"="ProcessExplicit",
         "Problem"="ProblemDefinitionExplicit",
         "Objectives"="ObjectivesExplicit",
         "Alternatives"="AlternativesExplicit",
         "Tradeoffs"="TradeOffsExplicit",
         "Decision"="DecisionExplicit",
         "Roles"="RolesExplicit",
         "Open Meetings"="OpenMeetings",
         "Adopted"="ResultsAdopted") %>%
  summarise_all(funs(sum))%>%
  gather(Explicit) %>%
  mutate(Percent=value/n_mse*100) %>%
  mutate(Explicit=factor(Explicit,levels=
                           c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision",
                             "Roles","Open Meetings","Adopted"))) %>%
  rename("Number"="value")

@@

# plot explicit documentation of steps or components of MSE processes
# Random Sample MSEs
Freq.plot<-ggplot(freq.data,aes(Explicit,Percent,fill=Analysis))+
  geom_col(position="dodge")+geom_vline(xintercept=3.5,linetype="dashed")+
  geom_vline(xintercept=1.5,linetype="dashed")+
  coord_flip()+scale_fill_manual(values=c("#66CCCC","#006666")) +
  scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
  scale_x_discrete(
    limits=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
             "Alternatives","Objectives","Problem","Process"),
    labels=c("Adopted","Open Meetings","Roles","Decision","Tradeoffs",
             "Alternatives","Objectives","Problem","Process")) +
  guides(fill = guide_legend(reverse=T)) + theme_bw() +
  theme(legend.position = c(0.85,0.07),legend.title = element_blank())

Part.plot<-ggplot(part.data,aes(x=order,y=Percent,fill=Analysis)) +
  facet_wrap(~Stage,scale="free",ncol=2) + geom_col(position="dodge") +
  scale_fill_manual(values=c("#66CCCC","#006666")) +
  scale_x_continuous(breaks = part.data$order,
                     labels = part.data$Participants)+
  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
  ylab("Percent")+xlab(NULL)+coord_flip() +
  guides(fill = guide_legend(reverse=T)) + theme_bw() +
  theme(legend.position = c(0.86,0.075),legend.title = element_blank())

@@

# Methods

# Results

>>>>

```{r load data, include=FALSE}
options(knitr.table.format = "html")
##### load libraries #####
library(tidyverse) # upgrade to base R
library(readxl) # read in excel files
library(shiny) # Shiny App
library(shinythemes) # web themes for Shiny
library(DT) # Data Table
library(purrr) # for loop alternative
library(ggmap) # map plotting
library(mapdata) # basemap generation
library(odbc) # database connection 
library(RMySQL) # MySQL scripting in R
library(DBI) # database interface
library(kableExtra) # extra formating of the tables

@@

# Move comment column to the end
data<-data %>%
  select(-'Comments', 'Comments')
```

```{r filter data, include=FALSE}

# Filter data by study type
data_pub<-filter(data,IncludeInPublication==TRUE)
data_CC<-filter(data,str_detect(Drivers,"Climate Change"))

@@

#####-- Filter Data by analysis --#####
# Get columns for study summary
summary.col<-c("Citation",
               "Species",
               "Location",
               "System")
# Get columns for study drivers and problem
prob.col<-c("Citation",
            "ProblemDefinition",
            "Drivers",
            "ConsequencePrediction",
            "TradeOffMethod_Exp",
            "TradeOffMethod_Sub",
            "Decision")
# Get columns for frequency analysis
freq.col<-c("ProcessExplicit",
            "ProblemDefinitionExplicit",
            "ObjectivesExplicit",
            "AlternativesExplicit",
            "TradeOffsExplicit",
            "DecisionExplicit")
freq2.col<-c("RolesExplicit",
            "OpenMeetings",
            "ResultsAdopted")
# Get columns for participant analysis
part.col<-c("Leader",
            "Participants",
            "ObjElicitationSource_Exp",
            "ProcedureElicitation_Exp",
            "ObjElicitationSource_Sub",
            "ProcedureElicitation_Sub")
obj.col<-c("ObjType",
           "ObjCategory",
           "ObjDirection",
           "ObjScale")
alt.col<-c("ManagementType",
           "AlternativesEvaluated")
# Get columns for map
map.col<-c("Latitude",
           "Longitude",
           "Citation",
           "Drivers")

# Select data summary columns
summary.data_pub<-data_pub %>%
    select(summary.col)
summary.data_CC<-data_CC %>%
  select(summary.col)
```

```{r analyze data, include=FALSE}
options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")

@@

freq.data_pub<-data_pub %>%
    select(freq.col) %>%
    rename("Process"="ProcessExplicit",
           "Problem"="ProblemDefinitionExplicit",
           "Objectives"="ObjectivesExplicit",
           "Alternatives"="AlternativesExplicit",
           "Tradeoffs"="TradeOffsExplicit",
           "Decision"="DecisionExplicit") %>%
    summarise_all(funs(sum))%>%
    gather(Explicit) %>%
    mutate(Percent=value/n_mse*100) %>%
    mutate(Explicit=factor(Explicit,levels=
                  c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision"))) %>%
    rename("Number"="value")

freq.data_CC<-data_CC %>%
  select(freq.col) %>%
  rename("Process"="ProcessExplicit",
         "Problem"="ProblemDefinitionExplicit",
         "Objectives"="ObjectivesExplicit",
         "Alternatives"="AlternativesExplicit",
         "Tradeoffs"="TradeOffsExplicit",
         "Decision"="DecisionExplicit") %>%
  summarise_all(funs(sum))%>%
  gather(Explicit) %>%
  mutate(Percent=value/n_mse*100) %>%
  mutate(Explicit=factor(Explicit,levels=
                           c("Process","Problem","Objectives","Alternatives","Tradeoffs","Decision"))) %>%
  rename("Number"="value")

@@

# plot explicit documentation of steps or components of MSE processes
# Random Sample MSEs
Freq.plot<-ggplot(freq.data_pub,aes(Explicit,Percent))+
  geom_col()+
  coord_flip() +
  scale_y_continuous(limits=c(0,100),expand = c(0,0))+xlab(NULL)+
  scale_x_discrete(
    limits=c("Decision","Tradeoffs",
             "Alternatives","Objectives","Problem","Process"),
    labels=c("Decision","Tradeoffs",
             "Alternatives","Objectives","Problem","Process")) + theme_bw()

Part.plot<-ggplot(part.data_pub,aes(x=order,y=Percent)) +
  facet_wrap(~Stage,scale="free",ncol=2) + geom_col() +
  scale_fill_manual(values=c("#66CCCC","#006666")) +
  scale_x_continuous(breaks = part.data$order,
                     labels = part.data$Participants)+
  scale_y_continuous(limits=c(0,100),expand = c(0,0)) +
  ylab("Percent")+xlab(NULL)+coord_flip() + theme_bw()

@@

### Objectives
We review published MSEs against the steps of the SDM framework to:

* Determine the rate of documentation, and use of steps, in the MSE process:  
* Identify participants in MSE processes.
* Highlight unique aspects of MSEs that model climate change as a driver.
* Identify aspects of successful MSEs that can be adopted to inform fisheries management in a changing climate  
* Provide insights about MSE processes and documentation that can further learning to improve future MSEs

# Methods
## Finding and Sampling the MSE Literature
We conducted our search for MSEs in the SCI-EXPANDED index from Web of Science, searching for “management strategy evaluation” by topic across all years on January 8th, 2019. This search returned 253 results.  We reviewed a random sample of 30 articles that document a MSE, removing articles that were reviews, meta-analyses, or simply cited other MSE articles from our sample. Of the 253 articles, 11 included climate change as a driver of fishery status, and after removing other articles we estimate 110-160 articles document a MSE.

| MSE type | Count |
|:--:|:--:|
| Published (estimated) | 110-160 |
| Random Sample | 30 |
| Climate Change | 11 |

## Reviewing MSE documentation 
Methods for scoring the documentation.

# Results
## Are structured decision making steps explicit in MSEs?
The majority of MSEs did not explicitly document how the stages of the MSE process -- defined using the SDM framework -- were completed (Table \@ref(tab:explicit)).  The most documented stage was the alternative stage, while 3 of the 30 studies explicitly documented a tradeoff analysis and 2 of the 30 documented a management decision.

```{r explicit}
freq.data_pub %>%
  mutate("Percent"=round(Percent,0)) %>% 
  kable(align = "c", caption = "Prevalence of Explicit Documentation of stages in the MSE process")
```

### Who is involved in MSEs?
The primary participants and leaders of the MSE processes were scientists, with some lead by governments and management agencies (Figure \@ref(fig:participants)).  Objectives and alternatives were elicited from a variety of participants, with scientists seemingly selecting these components in cases where documentation was inexplicit.

```{r participants, fig.cap='Who guided, participated in, or provided input during the specified steps of the MSE process', fig.height=6, out.width="100%"}
Part.plot
```

<!-- ### Where have climate change MSEs occurred? -->
<!-- To date MSEs modeling climate change have been concentrated in North America and Australia (Figure \@ref(fig:mapfigure)). -->

<!-- ```{r, mapfigure, fig.cap='Map of MSE study locations.', fig.height=4, out.width="100%"} -->
<!-- MSE.map + -->
<!--   theme(text = element_text(size=14)) -->
<!-- ``` -->

<!-- ### What questions and management objectives do MSEs address? -->
<!-- Uncertainty in system parameters was the main driver of the system state, but a variety of drivers were considered across the random sample of MSEs (Figure \@ref(fig:driversPlot)). -->

<!-- ```{r driversPlot, fig.cap='Drivers included in MSE models', fig.height=6, out.width="100%"} -->
<!-- Driver.plot + -->
<!--   theme(text = element_text(size=14)) -->
<!-- ``` -->

<!-- More objectives were considered per MSE in the random sample than in the climate change MSEs (Table \@ref(tab:PerMSE)).  Climate change MSEs were less likely to consider economic and social objectives than the random sample (Table \@ref(tab:objectiveCategories)). -->

<!-- ```{r, PerMSE} -->
<!-- color_Pub <- which(per.MSE$Analysis=='Random Sample') -->
<!-- color_CC <- which(per.MSE$Analysis=='Climate Change') -->

<!-- per.MSE %>% -->
<!--   kable(align = "c", caption = "Number of objectives considered per management strategy evaluation") %>%   -->
<!--   row_spec(color_CC, color = "#66CCCC", background = "white") %>% -->
<!--   row_spec(color_Pub, color = "#006666", background = "white") -->
<!-- ``` -->

<!-- ```{r, objectiveCategories} -->
<!-- color_Pub <- which(objcat.data$Analysis=='Random Sample') -->
<!-- color_CC <- which(objcat.data$Analysis=='Climate Change') -->

<!-- objcat.data %>% -->
<!--   kable(align = "c", caption = "Objective categories considered in management strategy evaluations") %>% -->
<!--   row_spec(color_CC, color = "#66CCCC", background = "white") %>% -->
<!--   row_spec(color_Pub, color = "#006666", background = "white") -->
<!-- ``` -->

### Main Conclusions
The picture painted by MSE documentation is one of scientist driven simulation studies rather than management driven decision processes. 
While documentation about the **modeling** conducted in a MSE was explicit, documentation about the **process used to make decisions** about the MSE model **were limited**. Scientists participated in, or seemingly directly selected the objectives and alternatives to evaluate in most MSEs. Facilitators and decision analysts -- participants focused on the decision making process -- were rarely participants in MSE processes.  Conservation, Economic, and Yield focused objectives were commonly considered, while **broader social objectives were rarely considered**.

Despite numerous applications of MSE, the applications that include climate change are limited, leaving uncertainty about the best approach to MSE use for fishereis management that accounts for a changing environment.

### Explore our review on the web!
Check out our shiny app (<https://jonathancummings.shinyapps.io/MSEreview/>).

### References
